---
title: "RWorksheet_Cababasay#3a"
output:
  pdf_document: default
  html_document: default
---

---
title: "RWorksheet_Cababasay#3a"
author: "Mike Jayson Cababasay"
output: pdf_document
---

# Worksheet 3a – Using Vectors, Matrices and Arrays in R

## 1. Using Vectors

### a. First 11 uppercase letters
```{r}
LETTERS[1:11]
```

### b. Odd-numbered letters
```{r}
LETTERS[seq(1, 26, by = 2)]
```

### c. Vowels in uppercase
```{r}
LETTERS[c(1,5,9,15,21)]
```

### d. Last 5 lowercase letters
```{r}
letters[22:26]
```

### e. Lowercase letters 15 to 24
```{r}
letters[15:24]
```

---

## 2. Average April Temperatures

### a. Character vector of cities
```{r}
city <- c("Tuguegarao City","Manila","Iloilo City","Tacloban","Samal Island","Davao City")
city
```

### b. Numeric vector of temperatures (°C)
```{r}
temp <- c(42,39,34,34,30,27)
temp
```

### c. Combine into data frame
```{r}
weather <- data.frame(city, temp)
weather
```

### d. Rename columns to City and Temperature
```{r}
names(weather) <- c("City","Temperature")
weather
```

### e. Show structure
```{r}
str(weather)
```

### f. Rows 3 and 4
```{r}
weather[c(3,4), ]
```

### g. Highest and lowest temperature cities
```{r}
weather[which.max(weather$Temperature), ]
weather[which.min(weather$Temperature), ]
```

---

## 3. Using Matrices

### a. Matrix from 1–8 and 11–14 (4 cols × 3 rows)
```{r}
mat <- matrix(c(1:8,11:14), nrow = 3, ncol = 4)
mat
```

### b. Multiply by 2
```{r}
mat2 <- mat * 2
mat2
```

### c. Row 2 content
```{r}
mat2[2, ]
```

### d. Columns 3–4 of rows 1–2
```{r}
mat2[1:2, 3:4]
```

### e. Columns 2–3 of row 3
```{r}
mat2[3, 2:3]
```

### f. Column 4 only
```{r}
mat2[, 4]
```

### g. Name rows and columns
```{r}
rownames(mat2) <- c("isa","dalawa","tatlo")
colnames(mat2) <- c("uno","dos","tres","quatro")
mat2
```

### h. Reshape original matrix to 2 cols × 6 rows
```{r}
dim(mat) <- c(6,2)
mat
```

---

## 4. Using Arrays

### a. Create 3-D array (4 cols × 2 rows)
```{r}
nums <- c(1,2,3,6,7,8,9,0,3,4,5,1)
arr <- array(rep(nums,2), dim = c(2,4,3))
arr
```

### b. Check dimensions
```{r}
dim(arr)
```

### c. Name rows, columns, and dimensions
```{r}
rownames(arr) <- letters[1:2]
colnames(arr) <- LETTERS[1:4]
dimnames(arr) <- list(
  row = letters[1:2],
  col = LETTERS[1:4],
  dim = c("1st-Dimensional Array","2nd-Dimensional Array","3rd-Dimensional Array")
)
arr
```
